A meta-analysis of fungicide effects on mungbean powdery mildew

### 2021-05-12 04_Severity_MetaAnalysis.Rmd
square transform severity and back transform estimates

Inspect disease severity for a normality.
```{r sev_hist}
hist(PM_dat_D$PM_final_severity)
```

@@

```{r Metafor-Sev-analysis, cache=TRUE}
PMsev_mv <- rma.mv(
   yi = PM_final_severity,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mv)
```

@@

contrast_Ssum %>%
   flextable()%>%
   autofit()
```

@@

```{r plotsevContrasts}
par(mar = c(5, 13, 4, 2) + 0.1)
plot(glht(PMsev_mv, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), yaxt = 'n')
axis(
   2,
   at = seq_along(contrast_Ssum$contrast),
   labels = rev(contrast_Ssum$contrast),
   las = 2,
   cex.axis = 0.8
)
```

@@

intercept <- round(PMsev_mv$b, 4)[1,1]

# create data.frame
results_mv <- data.frame(
   Moderator = c(
   "Intercept / No Spray control",
   "Early",
   "Late",
   "Late+",
   "Recommended",
   "Recommended+"
   ),
   N = k5,
   k = k6,
   Effect = round(PMsev_mv$b, 4),
   se = round(PMsev_mv$se, 4),
   CI_lower = round(PMsev_mv$ci.lb, 4),
   CI_upper = round(PMsev_mv$ci.ub, 4),
   z_val = round(PMsev_mv$zval, 4),
   p_val = reportP(PMsev_mv$pval, AsNumeric = FALSE, P_prefix = FALSE),
   eff_bar = round(1 -(intercept + PMsev_mv$b)/intercept,4)*100,
   eff_CI_high = round(1 -(intercept + PMsev_mv$ci.ub)/intercept,4)*100,
   eff_CI_low = round(1 -(intercept + PMsev_mv$ci.lb)/intercept,4)*100

@@

>>>>

Inspect disease severity for a normality.
```{r sev_hist}
hist(PM_dat_D$PM_final_severity)
summary(PM_dat_D$PM_final_severity)
```
This meta-analysis assumes the effect size to be normally distributed, however, we can see here that it is skewed towards the top of the scale, eight and nine in this case.

Traditionally disease severity is log transformed due to it being more likely to be skewed to the lower disease severity scores. In this case if we `log()` transformed the response it would increase the skew towards higher disease severity.

We can improve the normality by transforming the effect size, disease severity, by squaring. As variance is proportional to the mean we square the variance aswell
```{r}
hist(PM_dat_D$PM_final_severity^2)
summary(PM_dat_D$PM_final_severity^2)
```

```{r}
# dat1 <- escalc(measure = "MN", yi = PM_final_severity., vi = vi, ni = replicates, data = PM_dat_D, slab = paste0(trial, spray_management))

PM_dat_D$yi_square <- PM_dat_D$PM_final_severity ^2
PM_dat_D$vi_square <- (PM_dat_D$vi +1)^2
```
Note: Here I added 1 to the variance before it was squared. This is because squaring numbers below one makes the number smaller or remain zero. This does not change the meta-analysis estimates but does increase tao
<br>  
<br>  
<br>

@@

```{r Metafor-Sev-analysis, cache=TRUE}
PMsev_mvSQ <- rma.mv(
   yi = yi_square,
   V = vi_square,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mvSQ)
```

```{r Metafor-Sev-analysis, cache=TRUE}

PMsev_mv <- rma.mv(
   yi = PM_final_severity,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)
summary(PMsev_mv)
```

@@

contrast_Ssum[6:15,] %>%
   mutate(coefficients = -sqrt(abs(coefficients)),
          StdErr = sqrt(StdErr))%>%
   flextable()%>%
   autofit()
```

@@

```{r plotsevContrasts}
par(mar = c(5, 13, 4, 2) + 0.1)
plot(glht(PMsev_mvSQ, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), yaxt = 'n')
axis(
   2,
   at = seq_along(contrast_Ssum$contrast),
   labels = rev(contrast_Ssum$contrast),
   las = 2,
   cex.axis = 0.8
)
```

@@

intercept <- round(PMsev_mvSQ$b, 4)[1,1]

# create data.frame
results_mv <- data.frame(
   Moderator = c(
   "Intercept / No Spray control",
   "Early",
   "Late",
   "Late+",
   "Recommended",
   "Recommended+"
   ),
   N = k5,
   k = k6,
   Effect = round(sqrt(abs(PMsev_mvSQ$b)), 4),
   se = round(sqrt(abs(PMsev_mvSQ$se)), 4),
   CI_lower = round(sqrt(abs(PMsev_mvSQ$ci.lb)), 4),
   CI_upper = round(sqrt(abs(PMsev_mvSQ$ci.ub)), 4),
   z_val = round(sqrt(abs(PMsev_mvSQ$zval)), 4),
   p_val = reportP(PMsev_mvSQ$pval, AsNumeric = FALSE, P_prefix = FALSE),
   eff_bar = round(sqrt(abs(1 -(intercept + PMsev_mvSQ$b)/intercept)),4)*100,
   eff_CI_high = round(sqrt(abs(1 -(intercept + PMsev_mvSQ$ci.ub)/intercept)),4)*100,
   eff_CI_low = round(sqrt(abs(1 -(intercept + PMsev_mvSQ$ci.lb)/intercept)),4)*100

@@

# restore negative numbers
results_mv[2:6, c(4,6:8,10:12)] <- -results_mv[2:6, c(4,6:8,10:12)]
